---
title: "p8105_hw1_zg2582"
author: Zaixin Gao (zg2582)
output: github_document
date: "2026-09-25"
---

```{r}
library(tidyverse)
```

# Problem 1
## load penguins dataset

```{r}
data("penguins", package = "palmerpenguins")
names(penguins)
nrow(penguins)
ncol(penguins)
summary(penguins)
```

this dataset contains 344 rows and 8 columns. The variables are: species, island, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g, sex, year. The mean of flipper length is 200.9.

## scatterplot of flipper_length_mm vs. bill_length_mm

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + 
  geom_point()

ggsave("penguin_scatterplot.pdf")
```

There is a positive relationship overall, penguins with longer bills tend to have longer flippers.
And the three species form separate clusters.

# Problem 2

## create df
```{r}
set.seed(1)
problem2_df = tibble(
  samp = rnorm(10),
  samp_pos = samp > 0,
  vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  vec_factor = factor(c("low", "low", "low", "mid", "mid", "mid", 
                        "high", "high", "high", "high"))
)
```

