---
title: "p8105_hw1_zg2582"
author: Zaixin Gao (zg2582)
output: github_document
date: "2026-09-25"
---

```{r}
library(tidyverse)
```

# Problem 1
## load penguins dataset

```{r}
data("penguins", package = "palmerpenguins")
names(penguins)
nrow(penguins)
ncol(penguins)
summary(penguins)
```

this dataset contains 344 rows and 8 columns. The variables are: species, island, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g, sex, year. The mean of flipper length is 200.9.
please see more important values in message. 

## scatterplot of flipper_length_mm vs. bill_length_mm

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + 
  geom_point()

ggsave("penguin_scatterplot.pdf")
```

There is a positive relationship overall, penguins with longer bills tend to have longer flippers.
And the three species form separate clusters. Adelie has short bills and flippers, Chinstrap has long bills and medium flippers, Gentoo has long bills and the longest flippers. 2 penguins with missing values weren't plotted.

# Problem 2

## create df
```{r}
set.seed(1)
problem2_df = tibble(
  samp = rnorm(10),
  samp_pos = samp > 0,
  vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  vec_factor = factor(c("low", "low", "low", "mid", "mid", "mid", 
                        "high", "high", "high", "high"))
)
```

## get means
```{r}
summary(problem2_df)
```

samp is numeric, so I got a full numeric summary. The mean is 0.1322.
samp_pos is logical, so R just counts FALSE and TRUE. 
vec_char is character, so R only describe the text. There's no mean reported.
vec_factor is factor, and R counts how many values are in each level. There's no mean here either.

###cross check with mean()

```{r}
mean(pull(problem2_df, samp))
mean(pull(problem2_df, samp_pos))
mean(pull(problem2_df, vec_char))
mean(pull(problem2_df, vec_factor))
```

the mean of last two returned NA, matchs what i got from summary().

##convert variables from one type to another

```{r, eval = FALSE}
as.numeric(pull(problem2_df, samp_pos))
as.numeric(pull(problem2_df, vec_char))
as.numeric(pull(problem2_df, vec_factor))
```

What I observed: the logical mean works because R turns TRUE and FALSE into 1 and 0. 
The character mean fails because letters can't become numbers. The factor mean fails because, even though factors are stored as numbers, those codes are category labels.

